Evolution of pre-copulatory and post-copulatory inbreeding behaviour and associated polyandry

### 2016-11-02 ms_pre_post.Rmd
New empirical studies in the discussion concerning monandry and inbreeding avoidance; new general prediction at end of discussion

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice.

>>>>

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also naturally depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and monandry is common, evolution of post-copulatory inbreeding avoidance is especially unlikely. For example, females of the gregarious cockroach *Blatella germanica* (L.) avoid inbreeding and are typically monandrous [@Lihoreau2007]; while occurrence of post-copulatory inbreeding avoidance in *B. germanica* has not, to our knowledge, been tested, we predict that such inbreeding avoidance is highly unlikely. Similarly, females of the solitary parasitoid wasp *Venturia canescens* are monandrous, and appear to avoid mating with siblings [@Metzger2010; @Metzger2010a]. We suggest that post-copulatory inbreeding avoidance in *V. canescens* would be highly unlikely to evolve.

General predictions
-------------------------------------------------------------------

Some broad predictions are possible given the verbal theory that we have developed and the populations that we have simulated. For example, given that post-copulatory inbreeding avoidance is useless for monandrous females and most effective for polyandrous females, while pre-copulatory inbreeding avoidance is perhaps most effective for monandrous females and less effective for polyandrous females, it would be interesting to test whether or not occurrence of post-copulatory inbreeding avoidance covaries positively, and pre-copulatory inbreeding avoidance covaries negatively, with degree of polyandory. To address such a question, more work is needed to identify the degree to which females of different species engage in polyandrous and the degree to which females express both pre-copulatory and post-copulatory inbreeding avoidance.

<!--- Some general conclusions addressing 1) what future theory should be developed, and 2) what future empirical studies need to consider. For 2, need to be aware of both pre- and post-copulatory inbreeding avoidance, potentially, and that the two are not independent, necessarily; cannot find any general theory for what happens when you have this kind of strategy redundancy -- two ways of getting at the same outcome. --->
